---
title: "Perform network meta-analysis"
author: Loukia M. Spineli
date: "Latest version: `r Sys.Date()`"
source: vignettes/perform_network_meta-analysis.Rmd
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
params:
  EVAL: !r identical(Sys.getenv("NOT_CRAN"), "true")
vignette: >
  %\VignetteIndexEntry{Perform network meta-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(rnmamod)
```

## Introduction

This vignette illustrates how to perform a random-effects network meta-analysis 
using the minimum required argumentation of the `run_model` function. 
 
## Example on a binary outcome 

We will use the network on pharmacologic interventions for chronic obstructive 
pulmonary disease (COPD) from the systematic review of [Baker et al. (2009)]
(https://pubmed.ncbi.nlm.nih.gov/19637942/). 

This network comprises 21 trials comparing seven pharmacologic interventions 
with each other and placebo. Theexacerbation of COPD (harmful outcome) is the 
analysed binary outcome.

```{r, echo = FALSE, results = "hide", fig.width = 5, fig.height = 5, fig.align = "center"}
netplot(data = nma.baker2009, drug_names = c("placebo", "budesonide \n", "budesonide plus \n formoterol \n", "fluticasone \n", "fluticasone plus \n salmeterol \n", "formoterol", "salmeterol \n", "tiotropium \n"),
        adjust.figsizex = 1.35, adjust.figsizey = 1.15)
```


